---
title: "Tidy Tuesday (02-09-2021) - US Census Data"
author: "Tim Gordon"
date: "February 9, 2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache=TRUE, warning=FALSE,
                      message=FALSE, echo = TRUE, dpi=180,
                      fig.width = 9, fig.height = 6)
library(tidyverse)
library(skimr)
library(ggplot2)
library(tidytext)
library(dplyr)
library(janitor)
library(ggpubr)
theme_set(theme_grey())
```

## Load data for Plastic Pollution

```{r}
lifetime_earn <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/lifetime_earn.csv')
student_debt <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/student_debt.csv')
retirement <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/retirement.csv')
home_owner <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/home_owner.csv')
race_wealth <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/race_wealth.csv')
income_time <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_time.csv')
income_limits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_limits.csv')
income_aggregate <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_aggregate.csv')
income_distribution <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_distribution.csv')
income_mean <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_mean.csv')
```

### Looking at income changes by Class over the years

```{r income_class_change}
income_time <- income_time %>%
  clean_names()

p1 <- income_time %>%
  mutate(percentile = factor(percentile, levels= c('10th','50th', '90th')),
         income_family = round(income_family, 0)) %>%
  filter(year > 1990) %>%
  ggplot(aes(x=year, y=income_family, label=round(income_family, 2), color=percentile)) +
  geom_line(size=2) +
  geom_point(data=subset(income_time, year %in% c(2000, 2006, 2008, 2012, 2016)), color='black', size=3) +
  geom_label(data=subset(income_time, year %in% c(2000, 2006, 2008, 2012, 2016)), position=position_dodge(width=1.2), vjust=-.07, hjust=.3) +
  labs(x='Year',
       y='Income',
       title='Income Changes by Income Class',
       subtitle = 'Income was highest in 2000. The lowest percentile never recovered while the highest percentile expanded their wealth during the new century') +
  theme_grey() +
  scale_y_continuous() 
```

### Further analysis into this centuries income levels.

### Lower 10th percentile family income dropped by 14%. 
### 50 percentile barely made even, while the top 90th increased income by nearly 10%.

```{r income_century_pct_change}

p2 <- income_time %>%
  filter(year %in% c(2000, 2016)) %>%
  arrange(percentile) %>%
  mutate(century_cmp = lag(income_family)) %>%
  mutate(pct_change = ((income_family-century_cmp)/income_family)) %>%
  filter(year == 2016) %>%
  ggplot(aes(percentile, pct_change, label=round(pct_change, digits=2), fill=percentile)) +#filter(year == 2016) 
  geom_col() + 
  coord_flip() +
  scale_y_continuous(labels = scales::percent)  +
  theme(legend.position = "none") +
  labs(x= "Income Group",
       y="Pct Change (%)", 
       title = "Family Income Pct Change from 2000 to 2016",
       subtitle = "The lowest 10% of family income lost almost 15% of their income. The top 10% increased their income by 9% over this period.")

ggarrange(p1, p2, labels = c('A', 'B'),
          common.legend =TRUE, legend="bottom",
          ncol=1)
```

### Student debt analysis
###g# All races are trending with a higher pct of student loans in later years. 
#### Hispanic trend is slow follow. Without context, it is hard to make a inference on why it may be.

```{r student_debt}
student_debt <- student_debt %>%
  clean_names()

student_debt %>% skim()

student_debt %>%
  ggplot(aes(year, loan_debt_pct, color=race)) +
  geom_line(size=1.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(title= "Student Loan Debt for different race groups",
       subtitle= 'All grows debt % increase steadily over the years.')

```

### Home ownership
#### Home ownership percentages are more volatile for minorities

```{r home_owner}
home_owner <- home_owner %>%
  clean_names()

home_owner %>% skim()

home_owner %>%
  ggplot(aes(year, home_owner_pct, color=race)) +
  geom_line(size=1.8) +
  theme_grey() +
  scale_y_continuous(labels = scales::percent) +
  labs(x='Year',
       y='Homeowner_PCT',
       title = 'Home Ownership by Race',
       subtitle = 'Minority home-ownership is more volatile throughout the years. Also is on average lower than White People.')
  
```

```{r}
income_distribution <- income_distribution %>%
  clean_names()

income_distribution %>% skim()

income_distribution %>%
  select(race) %>%
  unique()

income_distribution %>%
  select(income_bracket) %>%
  unique()

income_distribution %>%
  group_by(race) %>% 
  summarise(earliest = min(year))

### TO DO - use alone, exlude combinations (not much difference and keep hispanic any)

income_dist %>%
  select(income_bracket) %>%
  unique()

income_dist %>% skim()

income_distribution %>%
  # mutate(income_bracket = factor(income_bracket, levels=brackets)) %>%
  ggplot(aes(x=year, y=income_mean, color=race)) +
  geom_line()

income_dist %>%
  # mutate(income_bracket = factor(income_bracket, levels=brackets)) %>%
  ggplot(aes(x=year, y=income_mean, color=race)) +
  geom_line()

```


