---
title: "Tidy Tuesday (02-09-2021) - US Census Data"
author: "Tim Gordon"
date: "February 9, 2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache=TRUE, warning=FALSE,
                      message=FALSE, echo = TRUE, dpi=180,
                      fig.width = 9, fig.height = 6)
library(tidyverse)
library(skimr)
library(ggplot2)
library(tidytext)
library(dplyr)
library(janitor)
library(ggpubr)
theme_set(theme_grey())
```

## Load data for Plastic Pollution

```{r}
student_debt <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/student_debt.csv')
home_owner <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/home_owner.csv')
income_time <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_time.csv')
income_distribution <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_distribution.csv')

```

### Looking at income changes by Class over the years

```{r income_class_change}
income_time <- income_time %>%
  clean_names()

p1 <- income_time %>%
  mutate(percentile = factor(percentile, levels= c('10th','50th', '90th')),
         income_family = round(income_family, 0)) %>%
  filter(year > 1990) %>%
  ggplot(aes(x=year, y=income_family, label=round(income_family, 2), color=percentile)) +
  geom_line(size=2) +
  geom_point(data=subset(income_time, year %in% c(2000, 2006, 2008, 2012, 2016)), color='black', size=3) +
  geom_label(data=subset(income_time, year %in% c(2000, 2006, 2008, 2012, 2016)), position=position_dodge(width=1.2), vjust=-.07, hjust=.3) +
  labs(x='Year',
       y='Income',
       title='Income Changes by Income Class',
       subtitle = 'Income was highest in 2000. The lowest percentile never recovered while the highest percentile expanded their wealth during the new century') +
  theme_grey() +
  scale_y_continuous() 
```

### Further analysis into this centuries income levels.

### Lower 10th percentile family income dropped by 14%. 
### 50 percentile barely made even, while the top 90th increased income by nearly 10%.

```{r income_century_pct_change}

p2 <- income_time %>%
  filter(year %in% c(2000, 2016)) %>%
  arrange(percentile) %>%
  mutate(century_cmp = lag(income_family)) %>%
  mutate(pct_change = ((income_family-century_cmp)/income_family)) %>%
  filter(year == 2016) %>%
  ggplot(aes(percentile, pct_change, label=round(pct_change, digits=2), fill=percentile)) +#filter(year == 2016) 
  geom_col() + 
  coord_flip() +
  scale_y_continuous(labels = scales::percent)  +
  theme(legend.position = "none") +
  labs(x= "Income Group",
       y="Pct Change (%)", 
       title = "Family Income Pct Change from 2000 to 2016",
       subtitle = "The lowest 10% of family income lost almost 15% of their income. The top 10% increased their income by 9% over this period.")

ggarrange(p1, p2, labels = c('A', 'B'),
          common.legend =TRUE, legend="bottom",
          ncol=1)
```

### Home ownership
#### Home ownership percentages are more volatile for minorities

```{r home_owner}
home_owner <- home_owner %>%
  clean_names()

home_owner %>% skim()

home_owner %>%
  ggplot(aes(year, home_owner_pct, color=race)) +
  geom_line(size=1.8) +
  theme_grey() +
  scale_y_continuous(labels = scales::percent) +
  labs(x='Year',
       y='Homeowner_PCT',
       title = 'Home Ownership by Race',
       subtitle = 'Minority home-ownership is more volatile throughout the years. Also is on average lower than White People.')
  
```

```{r race-income-dist}
income_distribution <- income_distribution %>%
  clean_names()

income_distribution %>% skim()

### Not all years match identify earliest group category to use
income_distribution %>%
  group_by(race) %>% 
  summarise(earliest = min(year))

income_distribution %>%
  filter(race %in% c("Asian Alone", "Black Alone", "Hispanic (Any Race)", "White Alone")) %>% 
  drop_na() %>%
  # mutate(income_bracket = factor(income_bracket, levels=brackets)) %>%
  ggplot(aes(x=year, y=income_mean, color=race)) +
  geom_line(size=1.2) + 
  labs(x="Year",
       y="Income Mean",
       title="Income Mean Changes by Race")

```

### Student debt analysis
###g# All races are trending with a higher pct of student loans in later years. 
#### Hispanic trend is slow follow. Without context, it is hard to make a inference on why it may be.

```{r student_debt}
student_debt <- student_debt %>%
  clean_names()

student_debt %>% skim()

student_debt %>%
  ggplot(aes(year, loan_debt_pct, color=race)) +
  geom_line(size=1.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(title= "Student Loan Debt for different race groups",
       subtitle= 'All grows debt % increase steadily over the years.')

```

## Model Relationship 
### Model linear relationship between year and race for student debt %

```{r model_debt}
library(tidymodels)

### set model
linear_spec <- linear_reg() %>% 
  set_mode("regression") %>%
  set_engine("lm")

### set recipe
debt_recipe <- 
  recipe(loan_debt_pct ~ year + race, data=student_debt) %>%
  step_dummy(all_nominal())

debt_recipe

debt_prep <- prep(debt_recipe) # prepare recipe

debt_prep

student_prep <- juice(debt_prep)
student_prep ## apply recipe


linear_fit <- linear_spec %>%
  fit(loan_debt_pct ~ year * (race_Hispanic+race_White), data=student_prep)

tidy(linear_fit) ### p-value below .05 for all but white race interactions

glance(linear_fit$fit) ### .84 adjusted r^2 is good for amount of variables 
```

```{r gen-new-points}
test <- expand_grid(race = c("Black", "White", "Hispanic"), 
                 year = 1990:2024)
```

```{r model-cmp}
test_prep <- bake(debt_prep, new_data= test)

augment(linear_fit$fit, newdata=test_prep) %>% 
    mutate(race =case_when(race_Hispanic == 0 & race_White == 0 ~ 'Black',
                         race_Hispanic == 1 ~ 'Hispanic',
                         TRUE ~ 'White')) %>%
  ggplot(aes(x=year, y=.fitted, color=race)) +
  geom_line(size=1.2) +
  geom_line(data=student_debt, aes(year, loan_debt_pct, color=race))+
  annotate("rect",xmin=2016, xmax=2024, ymin=-Inf, ymax=Inf, alpha=.25) +
  scale_y_continuous(labels = scales::percent) +
  labs(x="year",
       y="% of student debt", 
       title = "Student loan debt by race",
       subtitle = "Linear regression used to compare and forecast current debt percentages.")
```

