---
title: "Tidy Tuesday (01-19-2021) - Kenya Census"
author: "Tim Gordon"
date: "January 24, 2021"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache=TRUE, warning=FALSE,
                      message=FALSE, echo = TRUE, dpi=180,
                      fig.width = 9, fig.height = 6)
library(tidyverse)
library(skimr)
library(ggplot2)
library(tidytext)
library(dplyr)
```

## Load data for Kenya Census 

```{r load_data}
gender <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-19/gender.csv')
crops <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-19/crops.csv')
household <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-19/households.csv')
```

## Gender
### Gender is fairly balanced between all counties

```{r gender}
gender_df <- gender %>%
  filter(County != "Total") %>%
  select(!Total) %>%
  mutate(County = str_to_lower(County)) %>%
  pivot_longer(!County, names_to="gender", values_to="pop") %>%
  mutate(gender = factor(gender, levels = c("Male","Female", "Intersex")))

gender_df %>%
  ggplot(aes(x=pop, y=County, fill=gender)) +
  geom_bar(stat="identity", position="fill", colour="black") +
  scale_x_continuous() + 
  theme_bw() +
  labs(title="Gender distribution across different Kenyan Counties",
       x="Population Distribution",
       y="County")

```

## Crops dataset
### Data clean and initial look at type of produce across the counties


```{r crops}
crops_df <- crops %>%
  pivot_longer(!SubCounty, names_to="Crops", values_to="Count") %>%
  filter(SubCounty != "KENYA") %>%
  mutate(Count = replace_na(Count, 0)) %>%
  mutate(Type = case_when(
    Crops %in% c("Cashew Nut", "Coconut", "Macadamia") ~ "Nut",
    Crops %in% c("Coffee", "Tea", "Khat(Miraa)") ~ "Plant",
    Crops == "Farming" ~ "Farming",
    TRUE ~ "Fruit")) %>%
  mutate(SubCounty = str_to_lower(SubCounty))

crops_df %>%  
  ggplot(aes(x=SubCounty, y=Count, fill=Crops)) +
  geom_bar(stat="identity", color="black", alpha=.7)+
  scale_y_continuous() +
  facet_grid(Type ~ .) +
  theme_grey() + 
  theme(axis.text.x = element_text(angle = 60, hjust =.9)) +
  labs(title="Crop Coverage through Kenyan Counties",
       x="County",
       y="Count")
```

### Crops highlighted areas which produce a balance of the different type of crops

```{r crops_balanced}
crops_df %>%
  mutate(Type = factor(Type, levels = c('Nut', 'Plant', 'Farming', 'Fruit'))) %>%
  group_by(SubCounty, Type) %>%
  summarise(amount = sum(Count)) %>%
  mutate( area = factor(ifelse(SubCounty %in% c("meru","nyeri","embu", "murang'a"),"Highlighted","Normal"))) %>%
  ungroup() %>%
  mutate(SubCounty = str_to_lower(SubCounty)) %>%
  ggplot(aes(y=SubCounty, x=amount, fill=Type,
             alpha=area)) +
  geom_col() +
  scale_alpha_manual(name = "SubCounty", values=c(0.35, 1.0)) + 
  theme_gray() + 
  labs(title="Kenyan Counties type of crop",
       subtitle = "Highlighted counties with a better balance of all four types",
      x="Amount Produced by Population",
      y="County")
```

## Prepare kenya household

```{r}
household_df <- household %>% 
  filter(County != "Kenya") %>%
  mutate(County = str_to_lower(County)) %>%
  mutate(County = str_replace(County, 'city', '')) 

household_df %>%
  ggplot(aes(y=County, x=Population)) + 
  geom_col()
```

## Prep to join crops and households

```{r}
crops_merge <- crops_df %>%
  select(!Crops) %>% 
  group_by(SubCounty, Type) %>%
  summarise(Count = sum(Count)) %>% 
  ungroup() %>%
  pivot_wider(names_from = 'Type', values_from='Count')
```

### Merging data together for analysis on population impact

```{r}
pop_crops <- household_df %>% 
  left_join(crops_merge, by=c('County' = 'SubCounty')) %>%
  replace(., is.na(.), 0) %>%
  mutate(Total = (Farming + Fruit + Nut + Plant))
```



###  To be done. Align counties (specifically) in crops and households to measure why a county produces as it does 

```{r}
pop_crops %>% 
  mutate( area = factor(ifelse(County %in% c("meru","nyeri","embu", 'nairobi', "murang'a"),"Highlighted","Normal"))) %>%
  ggplot(aes(x=Population,y=Total, label=County, color=area)) +
  geom_label(label.size = .15) + 
  scale_x_continuous(name="Population", labels = scales::comma) +
  scale_y_continuous(name="Total", labels = scales::comma) + 
  theme_grey() +
  theme(legend.position = "none") + 
  labs(title='Kenyan Counties Production vs Population',
       subtitle='Counties of interest highlighted')
```



```{r}
pop_crops %>% 
  mutate(area = factor(ifelse(County %in% c("meru","nyeri","embu", 'nairobi', "murang'a"),"Highlighted","Normal"))) %>%
  ggplot(aes(x=Population,y=Total, label=County, fill=area)) +
  geom_label(colour='white', vjust='inward') + 
  scale_x_continuous(name="Population", labels = scales::comma) +
  scale_y_continuous(name="Total", labels = scales::comma) + 
  scale_fill_manual(name="area", values= c("#BB0000", "#006600")) +
  theme_dark() +
  theme(legend.position = "none") + 
  labs(title='Kenyan Counties Production vs Population',
       subtitle='Counties of interest highlighted (diverse crops or lack of crops)')
```

###

