---
title: "Tidy Tuesday (01-12-2021) - Tate Art Museum"
author: "Tim Gordon"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache=TRUE, warning=FALSE,
                      message=FALSE, echo = TRUE, dpi=180,
                      fig.width = 9, fig.height = 6)
library(tidyverse)
library(skimr)
library(ggplot2)
library(tidytuesdayR)
library(tidytext)
library(dplyr)
```

## Part 1: Load in dataset for Tidy Tuesday (01-12-2021). 

```{r data}
tidytuesdayR::last_tuesday()
tt_data <- tt_load("2021-01-12")
tt_artist <- tt_data$artists
tt_art <- tt_data$artwork

tt_df <- tt_artist %>% left_join(tt_art, by=c("id" = "artistId"))


```

### Analysis of variables in dataset. 
### Explore dataset


```{r eda}
### Who are the artists without genders (typically schools)
tt_df %>%
  filter(is.na(gender)) %>%
  select(name) %>%
  count(name, sort=TRUE)

tt_df <- tt_df %>% 
  mutate(cont1 = str_split(placeOfBirth, ", ", simplify=TRUE)[,1], 
         cont2= str_split(placeOfBirth, ", ", simplify=TRUE)[,2]) %>% 
  mutate(var = case_when(cont2=="" ~ cont1,
                      TRUE ~ cont2)) %>%
  mutate(country_origin = fct_lump(var, prop= .01)) %>%
  select(-c(cont1, cont2))

### What are the top 5 countries artists come from
summary(tt_df$country_origin)

```


### I was interested in plots regarding the gender of the artists present in the museum. 
### I am also graphed the top 5 countries of origin to visualize any changes over time. 


```{r plots}

### Does the Museum diversify the art collected as time goes on?
tt_df %>%
  filter(!is.na(country_origin) & year > 1850) %>%
  group_by(country_origin, decade = 10 * (year %/% 10)) %>%
  count() %>%
  mutate(country_origin = factor(country_origin, levels=c('Deutschland', 'France', 'United Kingdom', 'United States', 'Other'))) %>%
  ggplot(aes(x=decade, y=n, fill=country_origin)) +
    geom_col(position='dodge') +
    labs(title = "Art collection from different countries over decades", 
         x = "Decade",
         y = "Number of artworks")

tt_artist %>%
  filter(!is.na(name) & !is.na(gender) & !is.na(yearOfBirth)) %>% 
  group_by(gender, decade = 10 * (yearOfBirth %/% 10)) %>%
  tally() %>% 
  ungroup() %>%
  mutate(gender = factor(gender, levels = c('Female','Male'))) %>%
  ggplot(aes(x=decade, y=n, fill=gender)) +
    geom_bar(stat="identity", position="dodge", colour="black") +
    labs(title = 'Tate Artist Distribution across the decades',
         x="Decades (from 1500 to 2000)",
         y="Number of Artists in Tate")

tt_df %>%
  filter(!is.na(artist) & !is.na(gender) & !is.na(yearOfBirth) & artistRole=='artist' & acquisitionYear >= 1900) %>% 
  group_by(gender, decade = 10 * (acquisitionYear %/% 10)) %>%
  # mutate(gender = factor(gender, levels = c('Male','Female'))) %>%
  count() %>% 
  ungroup() %>%
  mutate(gender = factor(gender, levels = c('Female','Male'))) %>%
  ggplot(aes(x=factor(decade), y=n, fill=gender)) +
    geom_bar(stat='identity', position="dodge", alpha=.7) +
    theme(legend.position = "none") +
    labs(title = "A century of artwork acquired each decade by gender", 
         x="Decades (after 1900)",
         y="Number of acquisitions") 
      #facet_wrap(~ gender)
```


### The documentation on the site indicated that the accession number prefix referred to a different area of the museum collection.

```{r areas}
tt_art %>% 
  mutate(suffix = str_extract(accession_number, "(N|AR|T|[A-Z])")) %>% 
  group_by(suffix) %>%
  count()
```

### I was interested in the difference between words from earlier art works and more modern pieces in the title. 
### I also was interested in what the more common mediums of art include.

```{r}
tt_art %>%
  filter(year > 1800 & year < 1900) %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words) %>%
  count(word, sort=TRUE) %>%
  na.omit() %>%
  filter(n > 200) %>%
  ggplot(aes(x=n, y=reorder(word, n))) +
  geom_col()

tt_art %>%
  filter(year > 1900) %>%
  unnest_tokens(word, title) %>%
  anti_join(stop_words) %>%
  count(word, sort=TRUE) %>%
  na.omit() %>%
  filter(n > 200) %>%
  ggplot(aes(x=n, y=reorder(word, n))) +
  geom_col()


tt_art %>%
  unnest_tokens(word, medium) %>%
  anti_join(stop_words) %>%
  count(word, sort=TRUE) %>%
  na.omit() %>%
  filter(n > 500) %>%
  ggplot(aes(x=n, y=reorder(word, n))) +
  geom_col()
```

